---
title: "D3 Chord"
author: "Amar Sarvepalli"
date: "September 15, 2017"
output: html_document
---

# Debug:
For any given Origin or Destination, report mode-split between toll and non-toll,
  Total path variables: toll cost, travel time for each of the modes
  Segment path varaibles: Segments traversed, toll charged at each segment, v/c, speeds on each segment
  Link variables: Total v/c (back or queues) based on capacity.

# TODO:

## Graph: Shortest paths
#### Input:
 1. Edge File: A-B Links with distance, FTYPE, time as link variables.
 2. Vertex File: Node, type (zone, dmn, segment_id)
#### Logic:
 1. Build a shortest path between every zone to dmn
 2. Apply clustering or pick up closest 1 dmn for each taz (identifying segments)
    Note: clustering allows to quickly change when the shortest path changes
 
## Group: zones by segments
Group trip tables by segments allows to display eligible trip flows between the segments


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(chorddiag)
library(igraph)
library(dplyr)
library(tidyr)
library(knitr)

```


## User Settings
```{r}

# trip table input (in csv format)
# path <- "C:\\projects\\R-projects\\ELToD_Debugger\\data"
path <- "M:\\Projects\\Veterans ELToDv2.3 2017-0628_PDNE\\Base\\Y2020Rev\\PeakHour"
triptable_file <- "triptable_2020.csv"

# Setting to use one or two scenarios (old and new)
use_both <- TRUE 
use_new <- TRUE  # used only when use_both is false

# Working directory locations 
new_pp <- "M:\\Projects\\Veterans ELToDv2.3 2017-0628_PDNE\\Base\\Y2020Rev\\PeakHour\\Delay"
old_pp <- "M:\\Projects\\Veterans ELToDv2.3 2017-0628_PDNE\\Base\\Y2020Rev\\PeakHour\\NoDelay"

# Detailed OD travel time, toll costs to evaluate
# Selected Origins / Destinations
# SOV_file <- "RUNDETAIL_SOV17.PRN"
# otaz <- c(8,40, 9, 12, 13, 14) 
# dtaz <- c(33, 32)

# Detailed OD travel time, toll costs to evaluate
# Selected Origins / Destinations
SOV_file <- "RUNDETAIL_SOV8.PRN"
otaz <- c(31, 34)
dtaz <- c(41, 44)

```

## DEBUG: OD Trips
Writes out OD trips, EL Shares, and all utility attributes of the logit model for the selected OD pair(s). 
```{r}
# TODO: 
# Get all hours  and let user choose which hour(s) to report

# Just use data from the local directory for one file
if(!use_both) {
  
 if(!use_new) {
   # OLD: read OD outputs
   old_pp <- path
   old_df <- read.csv(paste(old_pp, SOV_file, sep = "\\"))
   max_iter_old <- max(old_df$ITER)
   debug_old <- old_df %>% filter(ITER == max_iter_old, I == otaz, J %in% dtaz)
   kable(debug_old, caption = "Select IJ No-Build") 
  
 } else {
   # NEW: read OD outputs
   new_pp <- path
   new_df <- read.csv(paste(new_pp, SOV_file, sep = "\\"))
   max_iter_new <- max(new_df$ITER)
   debug_new <- new_df %>% filter(ITER == max_iter_new, I %in% otaz, J %in% dtaz)
   # debug_new <- new_df %>% filter(ITER == max_iter_new, I %in% otaz) %>% arrange(desc(Trip_TOT))
   kable(debug_new, caption = "Select IJ Build")   
 }

}  else {
  # OLD: read OD outputs
  old_df <- read.csv(paste(old_pp, SOV_file, sep = "\\"))
  max_iter_old <- max(old_df$ITER)
  debug_old <- old_df %>% filter(ITER == max_iter_old, I == otaz, J %in% dtaz)
  kable(debug_old, caption = "Select IJ No-Build") 
  
  # NEW: read OD outputs
  new_pp <- path
  new_df <- read.csv(paste(new_pp, SOV_file, sep = "\\"))
  max_iter_new <- max(new_df$ITER)
  debug_new <- new_df %>% filter(ITER == max_iter_new, I == otaz, J %in% dtaz)
  kable(debug_new, caption = "Select IJ Build")   
  
}
  


```



```{r}
# Compute logit shares

```

# Gainers and Loosers
Write out top 10 OD pairs who gained / lost EL trips between No-Build ad Build scenarios

1. Checks OD trips between inputs (No-Build vs Build). Ideally these should 
    be the same input trip tables and reports if they are not identical.
2. Print top 10 ODs with EL trip gains
3. Print top 10 ODs with EL trip loses

```{r}

debug_old  <- old_df %>% 
              filter(ITER == max_iter_old) 

debug_new  <- new_df %>% 
              filter(ITER == max_iter_new)

debug_diff <- debug_new %>% 
              left_join(debug_old, by = c("I", "J"), suffix = c(".new", ".old")) %>%
              mutate(Trip_TOT.diff = Trip_TOT.new - Trip_TOT.old,
                     Trip_TL.diff = Trip_TL.new - Trip_TL.old,
                     Trip_TL_F.diff = Trip_TL_F.new - Trip_TL_F.old,
                     Trip_EL.diff = Trip_TL.diff + Trip_TL_F.diff)

# Trip tables with incosistent total OD flows
asymetry_IJ <- debug_diff %>% 
               filter(Trip_TOT.diff != 0) %>% 
               select(I, J, Trip_TOT.new, Trip_TOT.old, Trip_TOT.diff) %>%
               kable(caption = "Inconsistent Total OD Flows between No-Build and Build")

# Top 10 ODs with EL trip gains
top_10_gainers <- debug_diff %>% 
               filter(Trip_EL.diff > 1) %>% 
               select(I, J, Trip_TOT.new, Trip_TOT.old, 
                      Trip_TL.new, Trip_TL.old, 
                      Trip_TL_F.new, Trip_TL_F.old,
                      Trip_EL.diff) %>% 
               arrange(desc(Trip_EL.diff)) %>%
               slice(1:10) %>%
               kable(caption = "EL Trip Gains between No-Build and Build")

# Top 10 ODs with EL trip loses
top_10_loosers <- debug_diff %>% 
               filter(Trip_EL.diff < -1 ) %>% 
               select(I, J, Trip_TOT.new, Trip_TOT.old, 
                      Trip_TL.new, Trip_TL.old, 
                      Trip_TL_F.new, Trip_TL_F.old,
                      Trip_EL.diff) %>% 
               arrange(Trip_EL.diff) %>% 
               slice(1:10) %>%
               kable(caption = "EL Trip Gains between No-Build and Build")


# print tables
asymetry_IJ

top_10_gainers

top_10_loosers

```


# District to District Flows (Input OD trips)
For both input and output trips, generate aggegated d2d flows
1. Groups zones into districts based on the proximity (distance or time).
2. Aggregates input OD trips into these districts.
    b. Write out an aggregated trip table
    c. Produce a d3 plot (district to distric flows)
3. Aggregate output (No-Build and Build) EL trips into same districts
   a. Write out EL trips by districts
   b. Produce a d3 plot
   c. Produce a d3 difference plot to see gainers and loosers by segments.
   

```{r }
# Read trip table
tt <- read.csv(triptable_file)

# reformat as matrix
tt_m <- as.matrix(tt[, 2:ncol(tt)])
taz_names <- tt[,1]
dimnames(tt_m) <- list(otaz = taz_names,
                    dtaz = taz_names)

# produce a d3 chord diagram (group tazs by segments)
# chorddiag(tt_m, groupnamePadding = 80)

# filter (get the top trips for a given origin)
tt_t <- tt %>% gather(key, value, -SOV) 
  
tt_t %>% 
  filter(SOV == otaz) %>% 
  arrange(desc(value) )

```


